---
output:
rmarkdown::html_document:
    theme: lumen
    toc_depth: 5
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(knitr)
library(kableExtra)
library(pander)
```

# Using R for dNHC Research {.tabset .tabset-fade .tabset-pills}

## Setup

See 00\_Setup.R

### Check installed packages

Create a list of required packages

```{r eval=FALSE, include=TRUE}
list.of.packages <- c("dplyr", # v.0.8.5
                      "tidyr", # v.1.0.2
                      "plyr", # v.1.8.6
                      "spocc", # v.1.0.8
                      "ridigbio", # v.0.3.5
                      "tibble", # v.3.0.0
                      "rbison",
                      "CoordinateCleaner",
                      "lubridate",
                      "ggplot2",
                      "gtools",
                      "leaflet", 
                      "sp", 
                      "rgeos", 
                      "ggspatial", 
                      "raster", 
                      "agricolae", 
                      "gridExtra")

```

### Compare to installed packages

Compare list of required to installed packages. Retain only missing packages (!).

```{r eval=FALSE, include=TRUE}
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
```

### Install missing packages

Run for loop to install any missing packages. This loops through the list of new.packages and installs each.

```{r eval=FALSE, include=TRUE}
for(i in 1:length(new.packages)){
  install.packages(new.packages[i])
} 
```

## Cleaning Data

See 01_CleaningData.R     
Download data and clean it.    

### Load Packages

```{r message=FALSE, warning=FALSE}
library(CoordinateCleaner)
library(lubridate)
```

### Load Functions

This is a function I created with Natalie Patten (a very talent undergrad at UF). It will be part of her R package gatoRs (Geographic And Taxonomic Occurrence R-based Scrubbing).

```{r message=FALSE, warning=FALSE}
source("functions/DownloadingDataMore.R")
```

### Download Shortia galacifolia

Use spocc\_combined to create a csv containing occurrence records from iDigBio, GBIF, and BISON. Find out more about this function [here](https://github.com/mgaynor1/long-winded-scripts/tree/master/OccurrenceData).

```{r}
spocc_combine("Shortia galacifolia", "data/Shortia_galacifolia_raw.csv")
```

### Read in downloaded data frame

```{r}
rawdf <- read.csv("data/Shortia_galacifolia_raw.csv")
```

------------------------------------------------------------------------

### Cleaning

#### Inspect data frame

**What columns are included?**

```{r}
names(rawdf)
```

**How many observations do we start with?**

```{r}
nrow(rawdf)
```

#### <span style="color: blue;"> 1. Resolve taxon names </span>

Inspect scientific names included in the raw df.

```{r}
unique(rawdf$name)
```

Create a list of accepted names based on the name column in your data frame

```{r}
search <-  c("Shortia galacifolia")
```

Filter to only include accepted namesUsing the R package dplyr, we:
  1.  Filter the data frame to only include rows with the accepted names.
  2.  Filter out any rows with NA for dwc.scientificName.
  3.  Create a column called name and set it equal to "Shortia galacifolia".

```{r}
df <- rawdf %>% 
      filter(grepl(search, name, ignore.case =  TRUE)) %>% 
      mutate(new.name = "Shortia galacifolia")
```

**How many observations do we have now?**

```{r}
nrow(df)
```

#### <span style="color: blue;"> 2. Decrease number of columns </span>

##### Merge the two locality columns

```{r}
df$Latitude <-  dplyr::coalesce(df$Latitude, df$spocc.latitude)
df$Longitude <-  dplyr::coalesce(df$Longitude, df$spocc.longitude)
```

##### Merge the two date columns

```{r}
df$date <-  dplyr::coalesce(df$date, df$spocc.date)
```

##### Subset columns
Using the R package dplyr, we select and rename columns.   

```{r}
df <- df %>% 
      dplyr::select(ID = ID,
                name = new.name,
                basis = basis,
                coordinateUncertaintyInMeters = coordinateUncertaintyInMeters,
                informationWithheld = informationWithheld,
                lat = Latitude,
                long = Longitude,
                date = date)
```

#### <span style="color: blue;"> 3. Clean localities </span>

##### Remove NAs

Using the R package dplyr, we:
    1. Filter out(!) any rows where long 'is.na'.
    2. Filter out(!) any rows where lat 'is.na'.

```{r}
df <- df %>% 
      filter(!is.na(long)) %>% 
      filter(!is.na(lat))
```

**How many observations do we have now?**

```{r}
nrow(df)
```

##### Fix precision

Using the R base function 'round', we round lat and long to two decimal places.

```{r}
df$lat <- round(df$lat, digits = 2)
df$long <- round(df$long, digits = 2)
```

##### Remove unlikely points

Remove points at 0.00, 0.00 Using the R package dplyr, we:  
    1. Filter to retain rows where long is NOT(!) equal to 0.00.
    2. Filter to retain rows where long is NOT(!) equal to 0.00.

```{r}
df <- df %>% 
      filter(long != 0.00) %>% 
      filter(lat != 0.00)
```

##### Remove bad coordinates

Remove coordinates in cultivated zones, botanical gardens, and outside our desired range Using the R package CoordinateCleaner, we first if points are at biodiversity institutions and remove any points that are occurring at institutions.

```{r}
df <- cc_inst(df, 
              lon = "long",
              lat = "lat",
              species = "name")
```

Next, we look for geographic outliers and remove outliers.

```{r}
df <- cc_outl(df,
              lon = "long",
              lat = "lat",
              species = "name")
```

**How many observations do we have now?**

```{r}
nrow(df)
```

#### <span style="color: blue;"> 4. Remove Duplicates </span>

##### Fix dates

Using the R package lubridate, we first parse the date into the same format.

```{r}
df$date <- lubridate::ymd(df$date)
```

Next you are going to seperate date into year, month, and day - where every column only contains one set of information.

```{r}
df <- df %>% 
      dplyr::mutate(year = lubridate::year(date),
                month = lubridate::month(date),
                day = lubridate::day(date))
```

#### Remove rows with identical lat, long, year, month, and day

If a specimen shares lat, long, and event date we are assuming that it is identical. Many specimen lack date and lat/long, so this may be getting rid of information you would want to keep.

```{r}
df <- distinct(df, lat, long, year, month, day, .keep_all = TRUE)
```

**How many observations do we have now?**

```{r}
nrow(df)
```

#### <span style="color: blue;"> 5. Save Cleaned .csv </span>

```{r}
write.csv(df, "data/Shortia_galacifolia_061521-cleaned.csv", row.names = FALSE)
```

## Mapping

See 02_Mapping.R.    
Map occurrence records using ggplot2 and leaflet.       

### Load Packages

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(leaflet)
library(sp)
library(rgeos)
library(ggspatial)
```

### Read data file

```{r}
df <- read.csv("data/Shortia_galacifolia_061521-cleaned.csv")
```

### <span style="color: blue;"> Map using ggplot2 </span>
Make a simple map using ggplot2.

#### Set base maps  

Here we download a USA map, a state level USA map, and a county level USA map.

```{r}
USA <- borders(database = "usa", colour = "gray50", fill = "gray50")
states <-  borders("state", colour = "black", fill = NA)
county <- borders(database = "county", colour = "gray40", fill = NA)
```

#### Plot

Here we are going to add our USA map and the occurrence points. Next, we zoom into our area of interest. Then, we fix the x and y labels. Finally, we add a scale and a north arrow

```{r}
simple_map <- ggplot() +
              USA +
              # Better to add state or county here - order matters! 
              geom_point(df, 
                         mapping = aes(x = long, y = lat), 
                         col = "blue") +
              coord_sf(xlim = c(-86, -76), ylim = c(30, 38)) +
              xlab("Longitude") +
              ylab("Latitude") +
              annotation_scale() +
              annotation_north_arrow( height = unit(1, "cm"),
                                      width = unit(1, "cm"), 
                                      location = "tl")
```

```{r message=FALSE, warning=FALSE}
simple_map
```

##### Add States

```{r message=FALSE, warning=FALSE}
simple_map + states
```

##### Add Counties

```{r message=FALSE, warning=FALSE}
simple_map + county
```

### <span style="color: blue;"> Map using Leaflet </span>

Find out more about [leaflet here](https://rstudio.github.io/leaflet/). Basically it uses JavaScript to make interactive maps.   
   
#### Make points spatial     

```{r}
pts <- SpatialPoints(df[,7:6])
```

#### Make a simple leaflet map  

```{r}
m <- leaflet(pts) %>% 
     addCircles() %>% 
     addTiles() 
m
```

#### Set extent and zoom out

```{r}
m <- fitBounds(m, -86, 30, -76, 38)
m
```

#### Make fancy icon

```{r}
ShortiaIcon <- makeIcon(
  iconUrl = "https://live.staticflickr.com/2914/32392790054_27192e77e6_b.jpg",
  iconWidth = 18, iconHeight = 28,
  iconAnchorX = 17, iconAnchorY = 27
)

m <- leaflet(pts) %>% 
     addTiles()  %>% 
     addMarkers(icon = ShortiaIcon) 
m
```

#### Add label to see long, lat

```{r}
m <- leaflet(pts) %>% 
     addTiles()  %>% 
     addMarkers(label = paste0(df$long, ", ", df$lat))  
m 
```

#### Add rectangle marking extent

```{r}
m <- leaflet(pts) %>% 
     addTiles()  %>% 
     addMarkers(label = paste0(df$long, ", ", df$lat))   %>%
     addRectangles(
        lng1 = max(df$long), lat1 = max(df$lat),
        lng2 = min(df$long), lat2 = min(df$lat),
        fillColor = "transparent")
m
```

## Point Statistics

See 03_PointStats.R.     
Point sample and run a simple ANOVA.   

### Load Packages

```{r message=FALSE, warning=FALSE}
library(gtools)
library(raster)
library(dplyr)
library(tibble)
library(agricolae)
library(gridExtra)
```

### Read data files

```{r}
dfS <- read.csv("data/Shortia_galacifolia_061521-cleaned.csv")
dfG <- read.csv("data/Galax_urceolata_061521-cleaned.csv")
```

### Modify the data frame

#### Combined the two df

```{r}
df <- rbind(dfS, dfG)
```

#### Subset spatial points

```{r}
pts <- SpatialPoints(df[,7:6])
```

### Import BioClim files

#### List BioClim files

```{r}
list <- list.files("data/bio/", 
                    full.names = TRUE,
                    recursive = FALSE)
```

#### Order list using gtools

```{r}
list <- mixedsort(sort(list))
```

#### Load the rasters

```{r}
envtStack <- raster::stack(list)
```

### Point Sample

##### Extract value for each point

```{r}
ptExtracted <- raster::extract(envtStack, pts)
```

#### Convert to data frame

```{r}
ptExtracteddf <- as.data.frame(ptExtracted)
```

#### Add species name

```{r}
ptExtracteddf <- ptExtracteddf %>%
                 mutate(name = as.character(df$name))
```

### ANOVA

ANOVA or an analysis of variance. To run an ANOVA, you first set the linear model using lm. Next, the aov function will run the ANOVA.

```{r}
bio1aov <- aov(lm(ptExtracteddf[, 1] ~ name, data = ptExtracteddf))
bio1aov
```

#### Tukey HSD test

Here we will do a multiple comparisons of treatments by means of Tukey.

```{r}
b1 <- HSD.test(bio1aov, trt = "name", alpha = 0.05)  
b1
```

#### Separate the groups

```{r}
bio1_group <- b1$groups
bio1_group <- rownames_to_column(bio1_group, var = "name")
bio1_group <- bio1_group %>%
              dplyr::select(name, groups)
bio1_group
```

#### Subset from the original df

Here we subset the name and bio\_1 column from the original df.

```{r}
part <- ptExtracteddf %>%
        dplyr::select(name, bio_1)
```

#### Join the subset df to the bio1\_group data frame

```{r}
bio1pl <- left_join(part, bio1_group, by = "name")
```

#### Plot

```{r message=FALSE, warning=FALSE}
bio1_aov_plot <- ggplot(bio1pl, aes(x = name, y = bio_1)) +
                  geom_boxplot(aes(fill = groups)) +
                  geom_text(data = bio1_group, 
                            mapping = aes(x = name,
                                          y = 20, 
                                          label = groups), 
                            size = 5, inherit.aes = FALSE) +
                  theme(axis.text.x = element_text(angle = 90, 
                                                   size = 8, 
                                                   face = 'italic'))
bio1_aov_plot
```

#### Loop through all variables

First, set a list of the variable names and an empty list to store the resulting plots in.

```{r}
variablelist <- colnames(ptExtracteddf)[1:20]
plotlist <- c()
```

```{r}
for(i in 1:20){
      ### ANOVA 
      bioaov <- aov(lm(ptExtracteddf[, i] ~ name, data = ptExtracteddf))
      ### Tukey HSD test
      b <- HSD.test(bioaov, trt = "name", alpha = 0.05)    
      ### Separate the groups
      bio_group <- b$groups
      bio_group <- rownames_to_column(bio_group, var = "name")
      bio_group <- bio_group %>%
                   dplyr::select(name, groups)
      ### Subset from the original df
      part <- ptExtracteddf %>%
              dplyr::select(name, variablelist[i])
      #### Join the subset df to the bio_group data frame
      biopl <- left_join(part, bio_group, by = "name")
      #### Plot
      plotlist[[i]] <- ggplot(biopl, aes(x = name, y = biopl[,2])) +
                       geom_boxplot(aes(fill = groups)) +
                       geom_text(data = bio_group, 
                                mapping = aes(x = name,
                                              y = (max(na.omit(biopl[,2]))* 1.3), 
                                              label = groups), 
                                size = 5, inherit.aes = FALSE) +
                               scale_x_discrete(labels = c('G', 'S')) +
                       ggtitle(label  = paste(variablelist[i]))
      
}
```

```{r fig.height=20, fig.width=12, message=FALSE, warning=FALSE}
gridExtra::grid.arrange(grobs = plotlist)
```
